# Linear Regression {#linreg}

```{r linreg-setup, include=F}
advertising = read_csv(here("data", "Advertising.csv"))

```


You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

This chapter is about *linear regression*, a very simple approach for supervised learning. 
In particular, linear regression is a useful tool for predicting a quantitative response. 
Linear regression has been around for a long time and is the topic of innumerable textbooks. 
Though it may seem somewhat dull compared to some of the more modern statistical learning approaches described in later chapters of this book, linear regression is still a useful and widely used statistical learning method. 
Moreover it serves as a good jumping-off point for newer approaches: as we will see in later chapters, many fancy statistical learning approaches can be seen as generalizations or extensions of linear regression. 
Consequently, the importance of having a good understanding of linear regression before studying more complex learning methods cannot be overstated. 
In this chapter, we review some of the key ideas of the linear regression model, as well as the least squares approach that is most commonly used to fit this model.

Recall the `r data_colour("Advertising")` data from Chapter 2. 
Figure 2.1 displays `r data_colour("sales")` (in thousands of units) for a particular product as a function of advertising budgets (in thousands of dollars) for `r data_colour("TV")`, `r data_colour("radio")`, and `r data_colour("newspaper")` media. 
Suppose in our role as statistical consultants we are asked to suggest, on the basis of this data, a marketing plan for next year that will result in high product sales.
What information would be useful in order to provide such a recommendation? 
Here are a few important questions that we might seek to address:

1. *Is there a relationship between advertising budget and sales?*  
  Our first goal should be to determine whether the data provide evidence of an association between advertising expenditure and sales. 
  If the evidence is weak, then one might argue no money should be spent on advertising!

2. *How strong is the relationship between advertising budget and sales?*  
  Assuming there is a relationship between advertising and sales, we would like to know the strength of that relationship.
  In other words, given a certain advertising budget, can we predict sales with a high level of accuracy? 
  This would be a strong relationship. 
  Or is a prediction of sales based on advertising expenditure only slightly better than a random guess? 
  This would be a weak relationship.
 
3. *Which media contribute to sales?*  
  Do all three media - TV, radio, and newspaper - contribute to sales, or do only one or two of the media contribute? 
  To answer this question, we must find a way to separate out the individual effects of each medium when we have spent the mony on all three media.
 
4. *How accurately can we estimate the effect of each medium on sales?*  
  For every dollar spent on advertising in a particular medium, by what amount will sales increase? 
  How accurately can we predict this amount of increase?
  
5. *How accurately can we predict future sales?*  
  For every given level of television, radio, and newspaper sales, what is our prediction for sales, and what is the accuracy of this prediction?
  
6. *Is the relationship linear?*  
  If there is an approximately straight-line relationship between advertising expenditure in the various media and sales, then linear regression is an appropriate tool.
  If not, then it may still be possible to transform the predictor of the response so that linear regression can be used.
  
7. *Is there synergy among the advertising media?*  
  Perhaps spending \$50,000 on television advertising and \$50,000 on radio advertising results in more sales than allocating \$100,000 in either television or radio individually. 
  In marketing, this is known as a *synergy* effect, while in statistics it is called an *interaction* effect.
  
It turns out that linear regression can be used to answer each of these questions. 
We will first discuss all of these questions in a general context, and then return to them in this specific context in Section 3.4.

## Simple Linear Regression {#simple-lin-reg}

*Simple linear regression* lives up to its name: it is a very straightforward approach for predicting a quantitative response *Y* on the basis of a single predictor variable *X*. 
It assumes that there is an approximately linear relationship between *X* and *Y*. 
Mathematically we can write this linear relationship as

\begin{equation} 
  Y \approx \beta_0 + \beta_1X
  (\#eq:linmod)
\end{equation} 

You might read "$\approx$" as "*is approximately modeled as*". 
We will sometimes describe \@ref(eq:linmod) by saying we are *regressing* Y on X (or Y *onto* X). 
For example, *X* may represent `r data_colour("TV")` advertising and *Y* may represent `r data_colour("sales")`. 
Then we can regress `r data_colour("sales")` onto `r data_colour("TV")` by fitting the model:

$$\color{#B44C1C}{\textbf{sales}} = \beta_0+\beta_1\times\color{#B44C1C}{\textbf{TV}}$$

In Equation \@ref(eq:linmod), $\beta_0$ and $\beta_1$ are two unknown constants that represent the *intercept* and the *slope* terms in the linear model. 
Together, $\beta_0$ and $\beta_1$ are known as the model *coefficients* or *parameters*. 
Once we have used our training data to produce estimates $\hat{\beta_0}$ and $\hat{\beta_1}$ for our model parameters, we can predict future sales on the basis of a particular value of TV advertising by computing

\begin{equation} 
  \hat{y} = \hat{\beta_0} + \hat{\beta_1}x
  (\#eq:linest)
\end{equation} 

where $\hat{y}$ indicates a prediction of $Y$ on the basis of $X=x$. 
Here we use the *hat* symbol, ^ , to denote the estimated value of an unknown parameter or coefficient, or to denote the predicted value of the response.

### Estimating the Coefficients {#simple-coef-est}

In practice, $\beta_0$ and $\beta_1$ are unknown.
So before we can use \@ref(eq:linmod) to make predictions, we must use data to estimate the coefficients. Let

$$(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$$

represent *n* observation pairs, each of which consists of a measurement of *X* and a measurement of *Y*.
In the `r data_colour("Advertising")` example, this data set consists of the TV advertising budget and product sales in $n = 200$ different markets.
(Recall that the data are displayed in **Figure 2.1**.)
Our goal is to obtain coefficient estimates $\hat{\beta_0}$ and $\hat{\beta_1}$ such that the linear model \@ref(eq:linmod) fits the available data well—that is, so that $y_i \approx \hat{\beta_0}+\hat{\beta_1}x_i$ for $i = 1,...,n$.
In other words, we want to find an intercept b0 and a slope b1 such that the resulting line is as close as possible to the $n = 200$ data points.
There are a number of ways of measuring *closeness*.
However, by far the most common approach involves minimizing the *least squares* criterion, and we take that approach in this chapter.
Alternative approaches will be considered in Chapter 6.

```{r advertising-lm, fig.align = "center", fig.cap=paste("*For the", data_colour("Advertising"), "data, the least squares fit for the regression of", data_colour("sales"), "onto", data_colour("TV"), "is shown.  The fit is found by minimizing the sum of squared errors.  Each grey line segment represents an error, and the fit makes a compromise by averaging their squares.  In this case a linear fit captures the essence of the relationship, although it is somewhat deficient in the left of the plot.*"), fig.width= 12, fig.asp = 9/16}
# Fit linear model regressing Sales on TV
adv_lm = lm(sales~TV, advertising)

# Add the predicted values to the advertising dataframe
advertising %>% 
  mutate(predicted = adv_lm$fitted.values) %>% 
  ggplot(., aes(x=TV, y = sales))+
  geom_point(colour = "red")+
  geom_segment(aes(xend = TV, yend = predicted), colour = "grey40")+
  stat_smooth(method = "lm", se = F, fullrange = T)+
  scale_x_continuous(n.breaks = 6)+
  scale_y_continuous(breaks = c(5,10,15,20,25))+
  labs(y = "Sales")+
  theme_bw(base_size = 14, base_family = "Roboto")
  
```

Let $\hat{y_i} = \hat{\beta_0}+\hat{\beta_1}x_i$ be the prediction for $Y$ based on the *i*th value of $X$. 
Then $e_i = y_i - \hat{y_i}$ represents the *i*th *residual*—this is the difference between the *i*th observed response value and the *i*th response value that is predicted by our linear model.
We define the *residual sum of squares* (RSS) as
$$RSS = {e_1^2}+{e_2^2}+...+{e_n^2},$$
or equivalently as

\begin{equation} 
  RSS = (y_1 - \hat{\beta_0} - \hat{\beta_1}x_1)^2 + (y_2 - \hat{\beta_0} - \hat{\beta_1}x_2)^2 + ... + (y_n - \hat{\beta_0} - \hat{\beta_1}x_n)^2.
  (\#eq:rss)
\end{equation} 

The least squares approach chooses $\hat{\beta_0}$ and $\hat{\beta_1}$ to minimize the RSS.
Using some calculus, one can show that the minimizers are

\begin{equation} 
\begin{aligned}
  \hat{\beta_1} &= \frac{\sum_{i=1}^{n}(x_i - \bar{x})(y_i-\bar{y})}{\sum_{i=1}^{n}(x_i-\bar{x})^2},\\
  \hat{\beta_0} &= \bar{y}-\hat{\beta_1}\bar{x},
  \end{aligned}
  (\#eq:rssmin)
\end{equation} 

where $\bar{y} \equiv \frac{1}{n}\sum_{i=1}^{n}y_i$ and $\bar{x} \equiv \frac{1}{n}\sum_{i=1}^{n}x_i$ are the sample means.
In other words, \@ref(eq:rssmin) defines the *least squares coefficient estimates* for simple linear regression.

\@ref(fig:advertising-lm) displays the simple linear regression fit to the `r data_colour("Advertising")` data, where $\hat{\beta_0} = 7.03$ and $\hat{\beta_1} = 0.0475$.
In other words, according to this approximation, an additional \$1,000 spent on TV advertising is associated with selling approximately 47.5 additional units of the product. 
In **Figure 3.2**, we have computed RSS for a number of values of $\beta_0$ and $\beta_1$, using the advertising data with `r data_colour("sales")` as the response and `r data_colour("TV")` as the predictor.
In each plot, the red dot represents the pair of least squares estimates $(\hat{\beta_0}, \hat{\beta_1})$ given by \@ref(eq:rssmin). 
These values clearly minimize the RSS.



### Assessing the Accuracy of the Coefficient Estimates {#assess-coef-acc}



Recall from **Equation (2.1)** that we can assume that the *true* relationship between $X$ and $Y$ takes the form $Y = f(X)+\epsilon$ for some unknown function $f$, where $\epsilon$ is a mean-zero random error term.
If $f$ is to be approximated by a linear function, then we can write this relationship as

\begin{equation}
  Y = \beta_0+\beta_1X + \epsilon.
  (\#eq:lin-rel)
\end{equation} 

Here $\beta_0$ is the intercept term—that is, the expected value of $Y$ when $X=0$ and $\beta_1$ is the slope—the average increase in $Y$ associated with a one-unit increase in $X$.
The error term is a catch-all for what we miss with this simple model: the true relationship is probably not linear, there may be other variables that cause variation in $Y$, and there may be measurement error.
We typically assume that the error term is independent of $X$.
The model given by \ref(eq:lin-rel) defines the *population regression line*, which is the best linear approximation to the true relationship between $X$ and $Y$.^[The assumption of linearity is often a useful working model. 
However, despite what many textbooks might tell us, we seldom believe that the true relationship is linear.]
The least squares regression coefficient estimates \ref(eq:rssmin) characterize the *least squares line* \ref(eq:linest).
The left-hand panel of **Figure 3.3** displays these two lines in a simple simulated example.  We created 100 random Xs, and generated 100 corresponding Ys from the model

\begin{equation}
  Y = 2+3X+\epsilon
  (\#eq:sim-mod)
\end{equation}

where $\epsilon$ was generated from a normal distribution with mean zero.
