--- 
title: "An Introduction to Statistical Learning: with Applications in R"
author: "Gareth James, Daniela Witten, Trevor Hastie, & Robert Tibshirani"
date: "`r Sys.Date()`"
css: style.css
split_by: section
bibliography:
- book.bib
- packages.bib
description: This is a minimal example of using the bookdown package to write a book.
  The output format for this example is bookdown::gitbook.
documentclass: book
output:
  bookdown::gitbook:
    split_by: "section"
link-citations: yes
site: bookdown::bookdown_site
biblio-style: apalike
---

```{r setup, include=F}
library(tidyverse)
library(ISLR)
library(here)
library(extrafont)
knitr::opts_chunk$set(echo=F, warning=F, message=F)


data_colour <- function(x) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", "#B44C1C", x)
  } else if (knitr::is_html_output()) {
    sprintf("<strong><span style='font-family:monospace; color: %s;'>%s</span></strong>", "#B44C1C", 
      x)
  } else x
}

```


# Preface

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

I (not an author) am compiling this book for myself in order to learn both ISLR's material and how to use the **bookdown** package. 


