--- 
title: "An Introduction to Statistical Learning: with Applications in R"
author: "Gareth James, Daniela Witten, Trevor Hastie, & Robert Tibshirani"
date: "`r Sys.Date()`"
css: style.css
split_by: section
bibliography:
- book.bib
description: This is a minimal example of using the bookdown package to write a book.
  The output format for this example is bookdown::gitbook.
documentclass: book
link-citations: yes
github-repo: "MokeEire/BookdownISL"
site: bookdown::bookdown_site
biblio-style: apalike
---

```{r setup, include=F}
library(extrafont)
library(tidyverse)
library(GGally)
library(ISLR)
library(here)
library(gt)
library(patchwork)

knitr::opts_chunk$set(echo=T, warning=F, message=F)


data_colour <- function(x) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", "#B44C1C", x)
  } else if (knitr::is_html_output()) {
    sprintf("<strong><span style='font-family:monospace; color: %s;'>%s</span></strong>", "#B44C1C", 
      x)
  } else x
}


theme_islr = function(){ 

    theme_bw(base_size = 14, base_family = "Roboto") %+replace%    #replace elements we want to change
    
    theme(
      
      #grid elements
      panel.grid.major = element_blank(),    #strip major gridlines
      panel.grid.minor = element_blank(),    #strip minor gridlines

      #since theme_minimal() already strips axis lines, 
      #we don't need to do that again
      
      #text elements
      plot.title = element_text(             #title
                   face = 'bold',            #bold typeface
                   hjust = 0,                #left align
                   vjust = 2),               #raise slightly
      
      plot.caption = element_text(           #caption
                   hjust = 1),               #right align
      

      axis.text.x = element_text(            #margin for axis text
                    margin=margin(5, b = 10))
      
      #since the legend often requires manual tweaking 
      #based on plot content, don't define it here
    )
}

reg_table = function(reg_model){
   summary(reg_model)$coefficients %>% 
    as_tibble(rownames = "term") %>% 
    rename(Coefficient = Estimate, 
           `t-statistic` = `t value`, 
           pval = `Pr(>|t|)`) %>% 
    mutate(term = str_replace_all(term, c("\\:" = "<em> X </em>",
                                          "I(?=\\()|[\\(\\)]" = "")),
           term = str_c("<span style='color:#B44C1C'>", term, "</span>"),
           `p-value` = case_when(pval < .0001 ~ "<0.0001",
                                 T ~ str_c(round(pval, 4)))) %>%
    select(-pval) %>% 
    gt() %>% 
    cols_label(term ="") %>% 
    cols_align(columns = vars(`p-value`), align = "right") %>% 
    fmt_markdown(columns = vars(term)) %>% 
    fmt_number(columns = vars(Coefficient),decimals = 4) %>% 
    fmt_number(columns = vars(`Std. Error`), decimals = 3) %>% 
    fmt_number(columns = vars(`t-statistic`), decimals = 2) %>% 
    tab_style(
        style = cell_borders(
            sides = "right",
            style = "solid",
            weight = px(2)
        ),
        locations = list(
            cells_body(
                columns = vars(term)
            ),
            cells_column_labels(
                columns = vars(term)
            )
        )
    ) %>% 
    tab_style(
        style = cell_borders(
            sides = c("bottom", "top"),
            weight = NULL
        ),
        locations = cells_body(
            columns = everything()
        )
    ) %>% 
  tab_options(
        column_labels.border.bottom.color = "black",
        column_labels.border.bottom.style = "solid",
        column_labels.border.top.color = "black",
        column_labels.border.top.style = "solid",
        table_body.border.bottom.color = "black",
        table_body.border.bottom.style = "solid",table.font.color = "black",
        row.striping.include_table_body = F)
}
```

# Preface

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

I (not an author) am compiling this book for myself in order to learn both ISLR's material and how to use the **bookdown** package. 


